---
title: "R Notebook"
output: html_notebook
---

#Load Packages
```{r}
library(lehdr)
library(dplyr)
```

#Usage
## Pull the Oregon (state = "or") 2014 (year = 2014), origin-destination (lodes_type = "od"), all jobs including private primary, secondary, and Federal (job_type = "JT01"), all jobs across ages, earnings, and industry (segment = "S000"), aggregated at the Census Tract level rather than the default Census Block (agg_geo = "tract").
```{r}
or_od <- grab_lodes(state = "or", year = 2014, lodes_type = "od", job_type = "JT01", 
           segment = "S000", state_part = "main", agg_geo = "tract")

head(or_od)
```
##pulls the Oregon AND Rhode Island (state = c("or", "ri")) for 2013 and 2014 (year = c(2013, 2014) or year = 2013:2014)
```{r}
or_ri_od <- grab_lodes(state = c("or", "ri"), year = c(2013, 2014), lodes_type = "od", job_type = "JT01", 
           segment = "S000", state_part = "main", agg_geo = "tract")     

head(or_ri_od)       
```
#Additional Examples
##Adding at County level signifiers
```{r}
md_rac <- grab_lodes(state = "md", year = 2015, lodes_type = "wac", job_type = "JT01", segment = "S000")

head(md_rac)

md_rac_county <- md_rac %>% mutate(w_county_fips = str_sub(w_geocode, 1, 5))

head(md_rac_county)
```

##Aggregating at the County level
```{r}
md_rac_county <- md_rac %>% mutate(w_county_fips = str_sub(w_geocode, 1, 5)) %>% 
  select(-"w_geocode") %>%
  group_by(w_county_fips, state, year, createdate) %>% 
  summarise_if(is.numeric, sum)

head(md_rac_county)

md_rac_county <- grab_lodes(state = "md", year = 2015, lodes_type = "rac", job_type = "JT01", 
           segment = "S000", agg_geo = "county")
           
head(md_rac_county)
```

##Aggregating Origin-Destination
```{r}
md_od_county <- grab_lodes(state = "md", year = 2015, lodes_type = "od", job_type = "JT01", 
           segment = "S000", agg_geo = "county", state_part = "main")
           
head(md_od_county)
```

##Aggregating at Block Group, Tract, or State level
```{r}
md_rac_bg <- grab_lodes(state = "md", year = 2015, lodes_type = "rac", job_type = "JT01", 
           segment = "S000", agg_geo = "bg")
           
head(md_rac_bg)

md_rac_tract <- grab_lodes(state = "md", year = 2015, lodes_type = "rac", job_type = "JT01", 
           segment = "S000", agg_geo = "tract")
           
head(md_rac_tract)

md_rac_state <- grab_lodes(state = "md", year = 2015, lodes_type = "rac", job_type = "JT01", 
           segment = "S000", agg_geo = "state")
           
head(md_rac_state)
```