---
title: "bike_problem_work_sarah"
author: "Sarah Kohls"
date: "2022-11-18"
output: html_document
---

11/17/22 Work on bike problem

Load packages
```{r}
library(knitr)
library(tidyverse)
library(janitor)
library(lubridate) 
library(here) 
library(sf)
library(tmap)
library(tidycensus)
library(dsbox)
library(ggplot2)
library(gbfs)
library(dplyr)
library(stringr)
library(nabor)
```


```{r}
bikeshare = read_csv(here("202209-capitalbikeshare-tripdata.csv")) %>%
  clean_names() %>% select(ride_id, start_lat, start_lng, end_lat, end_lng) %>%
  filter(!is.na(end_lat), !is.na(end_lng))

census_api_key("74bc918a78c720321d756d915a6094c8b101b31f")

census=get_acs(geography = "tract",
               variables=c("pop"="B01001_001"),
               state="DC",geometry=TRUE,year=2018)
```

Count number of rides starting and ending in each census tract:
```{r}
census_wide = census %>% select(-moe) %>% spread(variable, estimate) %>%
  select(GEOID, pop, geometry) %>% st_transform(4326)

bikeshare_start_pts = bikeshare %>% select(start_lng, start_lat) %>%
  st_as_sf(coords=c("start_lng", "start_lat"), crs=st_crs(census_wide))

bike_census = census_wide %>% mutate(start_count = 
                                       lengths(st_intersects(., bikeshare_start_pts)))

bikeshare_end_pts = bikeshare %>% select(end_lng, end_lat) %>%
  st_as_sf(coords=c("end_lng", "end_lat"), crs=st_crs(census_wide))

bike_census = bike_census %>% mutate(end_count = 
                                       lengths(st_intersects(., bikeshare_start_pts)))
```

Plot:
```{r}
tmap_mode("view")
tm_shape(bike_census) + tm_polygons(c("pop", "start_count", "end_count"), alpha=0.5)
```

Normalized for population:
```{r}
bike_census_pop = bike_census %>% mutate(start_rate = start_count / pop,
                                         end_rate = end_count / pop)

tm_shape(bike_census_pop) + tm_polygons(c("start_rate", "end_rate"), alpha=0.5)
```

Filtered out huge popular region b/c messed up visualization:
```{r}
bike_census_pop = bike_census %>% mutate(start_rate = start_count / pop,
                                         end_rate = end_count / pop) %>%
  filter(GEOID != 11001006202)

tm_shape(bike_census_pop) + tm_polygons(c("start_rate", "end_rate"), alpha=0.5)
```

11/17/22 - Following Skufca's example from class:

Make data easily readable
```{r}
neigh=st_read(here("DC_Health_Planning_Neighborhoods.geojson")) %>% clean_names() %>%
  mutate(shapearea=st_area(geometry))
class(neigh)
```

Read basic Sep2022 data
```{r}
bike_data <- read_csv("202209-capitalbikeshare-tripdata.csv") %>% clean_names()
```

Filter to rides on Sep 1st to give a smaller dataset
```{r}
bike_data_use <- bike_data %>% 
  separate(started_at, c("date", "time"), " ") %>%
  separate(time, c("hour","minute","second"), ":") %>%
  mutate(minu = as.numeric(hour)*60 + as.numeric(minute))%>%
  separate(date, c("year","month","day"), "-") %>%
  filter(day == "01")
```

Create an sf object from the bikedata
```{r}
bike_data_sf=st_as_sf(bike_data_use, coords = c("start_lng", "start_lat"),
                      crs = 4326)
```

Join datasets
```{r}
neigh_temp=st_join(bike_data_sf,neigh,join = st_within) %>% group_by(code) %>%
  summarise(n=n()) %>%
  as_tibble() %>% select(-geometry)

neigh2=neigh %>% left_join(neigh_temp)
```

```{r}
tmap_mode("view")

tm_shape(neigh2) + tm_polygons("n", alpha = 0.5, n = 4, style = "quantile")
```

Joining with other spatial data
```{r}
census_api_key("74bc918a78c720321d756d915a6094c8b101b31f")

v20 = load_variables(2020,"acs5",cache=TRUE)
```

Get some data, old code:
```{r}
df_cencus=get_acs(geography = "tract",
                  variables=c("median_inc"="B06011_001",
                              "pop"="B01001_001",
                              "pop_black"="B02009_001"),
                  state="DC",geometry=TRUE,year=2018)
```

New code, trying to get at daytime population:
```{r}
df_cencus=get_acs(geography = "tract",
                  variables=c(
                    "total_pop"="B01003_001",
                    "worker_working"="08406_001",
                    "worker_living"="B08301_001"),
                  state="DC",geometry=TRUE,year=2020)
```

```{r}
dfc2=get_acs(geography = "tract",
             table="B08406",
             state="DC",geometry=TRUE,year=2020)
```

```{r}
class(df_cencus)
plot(df_cencus)
```

It's in long format -> make it wide
```{r}
df_cens=df_cencus %>% select(-moe) %>% spread(variable,estimate)

plot(df_cens)

tm_shape(df_cens) + tm_polygons("median_inc",alpha=0.5)
```



Homework for Monday 11/21:

Question to answer:
- Is peak in bike data actually people going to work or tourists?
- Are casual riders in the more touristy areas?
Look at open data dc site, tourist data, could use reverse geo coding to find location of monuments




