---
title: "bike_rentals"
author: "Sarah Kohls"
output: html_document
date: "2022-10-26"
---
##Load Packages
```{r}
library(tidyverse)
library(tidymodels)
library(dsbox)
```

##Recode season variable to be a factor
```{r}
dcbikeshare <- dcbikeshare %>%
  mutate(
    season = case_when(
      season == 1 ~ "winter",
      season == 2 ~ "spring",
      season == 3 ~ "summer",
      season == 4 ~ "fall"
      ),
    season = fct_relevel(season, "spring", "summer", "fall", "winter")
  )
```

##Recode holiday and workingday to be factors
```{r}
dcbikeshare <- dcbikeshare %>%
  mutate(
    holiday = ifelse(holiday == 0, "no", "yes"),      
    holiday = fct_relevel(holiday, "no", "yes"),    
    workingday = ifelse(workingday == 0, "no", "yes"),
    workingday = fct_relevel(workingday, "no", "yes")
  )
```

##Recode year variable to be a factor
```{r}
dcbikeshare <- dcbikeshare %>%
  mutate(
    yr = ifelse(yr == 0, "2011", "2012"),
    yr = fct_relevel(yr, "2011", "2012")
  )
```

##Recode weathersit variable
```{r}
dcbikeshare <- dcbikeshare %>%
  mutate(
    weathersit = case_when(
      weathersit == 1 ~ "clear",
      weathersit == 2 ~ "mist",
      weathersit == 3 ~ "light precipitation",
      weathersit == 4 ~ "heavy precipitation"
    ),
    weathersit = fct_relevel(weathersit, "clear", "mist", "light precipitation", "heavy precipitation")
  )
```

##Create raw values for temperature, humidity, and windspeed
```{r}
dcbikeshare <- dcbikeshare %>%
  mutate(
    temperature_raw = temp * 41,
    feeling_temperature_raw = atemp * 50,
    humidity_raw = hum * 100,
    windspeed_raw = windspeed * 67
  )
```

##Check cnt variable
```{r}
dcbikeshare %>%
  mutate(cas_plus_reg = casual + registered) %>%
  summarise(all_zero = all(cas_plus_reg == cnt))
```

##Visualization
```{r}
dcbikeshare %>%
  ggplot(mapping = aes(x = dteday, y = cnt, color = feeling_temperature_raw)) +
    geom_point(alpha = 0.7) +
    labs(
      title = "Bike rentals in DC, 2011 and 2012",
      subtitle = "Warmer temperatures associated with more bike rentals",
      x = "Date",
      y = "Bike rentals",
      color = "Temperature (C)"
    ) +
  theme_minimal()
```

##Bike rentals vs temperature
```{r}
cnt_tmp <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ temperature_raw, data = dcbikeshare)
cnt_tmp %>%
  tidy()
```
##Slope of model is approximately 162.
##For each additional degree celcius, we can expect bike rentals to increase by 162.
##The model has the intercept at approximately 1215.
##At a day w/ temp 0 degrees celcius, there will be about 1215 bike rentals. 

##R squared of model
```{r}
glance(cnt_tmp)$r.squared
```
##R-squared meaning: 39.4% of variability in number for total daily bike rental is explained by temperature.

##Bike rentals vs feeling temperature
```{r}
cnt_atmp <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ feeling_temperature_raw, data = dcbikeshare)
cnt_atmp %>%
  tidy()
```
##Model has intercept at about 946.
##Therefore on a day with feeling temp of 0 degrees celcius there will be about 946 bike rentals.
##Slope of model is about 150.
##Therefore for each additional degree in feeling temp, bike rentals will increase by about 150.

##R-sqaured of the model
```{r}
glance(cnt_atmp)$r.squared
```
##R-squared meaning: 39.8% of variability in number of total number of daily bike rentals is explained by feeling temperature.

##Feeling temperature is the better predictor since R-squared value is higher. 

```{r}
dcbikeshare
```


##Full model
```{r}
cnt_full <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season + yr + holiday + workingday + weathersit +
        temperature_raw + feeling_temperature_raw + humidity_raw + 
        windspeed_raw + feeling_temperature_raw * holiday, 
      data = dcbikeshare)
tidy(cnt_full)
```

##R-squared of the model
```{r}
glance(cnt_full)$adj.r.squared
```
##Intercept for non-holidays is higher than for holidays.
##Slope for feeling temp is steeper for holidays than for non-holidays

##Bike rentals highest in Fall, lowest in Winter.

##Backward selection

##Model w/out season
```{r}
rm_season <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ yr + holiday + workingday + weathersit +
        temperature_raw + feeling_temperature_raw + humidity_raw +
        windspeed_raw + feeling_temperature_raw * holiday, 
      data = dcbikeshare)
glance(rm_season)$adj.r.squared
```

##Model w/out year
```{r}
rm_year <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season + holiday + workingday + weathersit +
        temperature_raw + feeling_temperature_raw + humidity_raw +
        windspeed_raw + feeling_temperature_raw * holiday, data = dcbikeshare)
glance(rm_year)$adj.r.squared
```

##Model w/out holiday variable, removed interaction effect w/ holiday & feeling_temp_raw
```{r}
rm_holiday <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season + yr + workingday + weathersit +
        temperature_raw + feeling_temperature_raw + humidity_raw +
        windspeed_raw + feeling_temperature_raw, data = dcbikeshare)
glance(rm_holiday)$adj.r.squared
```

##Model w/out workingday variable
```{r}
rm_workingday <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season + yr + holiday + weathersit +
        temperature_raw + feeling_temperature_raw + humidity_raw +
        windspeed_raw + feeling_temperature_raw * holiday, data = dcbikeshare)
glance(rm_workingday)$adj.r.squared
```

##Model w/out weatersit variable
```{r}
rm_weathersit<- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season + workingday + yr + holiday +
        temperature_raw + feeling_temperature_raw + humidity_raw +
        windspeed_raw + feeling_temperature_raw * holiday, data = dcbikeshare)
glance(rm_weathersit)$adj.r.squared
```

##Model w/out temperature_raw variable
```{r}
rm_temp_raw <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season + weathersit + workingday + yr + holiday + feeling_temperature_raw + humidity_raw +
        windspeed_raw + feeling_temperature_raw * holiday, data = dcbikeshare)
glance(rm_temp_raw)$adj.r.squared
```

##Model w/out feeling_temperature_raw & removed interaction effect between holiday and feeling_temperature_raw
```{r}
rm_feeltemp_raw <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season + temperature_raw + weathersit + workingday + yr + holiday + humidity_raw +
        windspeed_raw, data = dcbikeshare)
glance(rm_feeltemp_raw)$adj.r.squared
```

##Model w/out humidity_raw variable
```{r}
rm_humidity_raw <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season + temperature_raw + weathersit + workingday + yr + holiday + feeling_temperature_raw +
        windspeed_raw + feeling_temperature_raw * holiday, data = dcbikeshare)
glance(rm_humidity_raw)$adj.r.squared
```

##Model w/out windspeed_raw variable
```{r}
rm_windspeed_raw <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season + humidity_raw + temperature_raw + weathersit + workingday + yr + holiday + feeling_temperature_raw + feeling_temperature_raw * holiday, data = dcbikeshare)
glance(rm_windspeed_raw)$adj.r.squared
```

##Model w/out interaction effect between holiday & feeling_temperature_raw
```{r}
rm_interaction <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season + windspeed_raw + humidity_raw + temperature_raw + weathersit + 
  workingday + yr + holiday + feeling_temperature_raw, data = dcbikeshare)
glance(rm_interaction)$adj.r.squared
```

##Conclusion: None of the modified models gives any improvement over the original full model
