---
title: "LAB 4 ML"
author: "Michael Lanfear"
date: "2022-10-17"
output: html_document
---
### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(devtools)
library(dplyr)
library(dsbox) 
```
```{r load-data, message = FALSE}
states <- read_csv("states.csv")
```

### Exercise 1
```{r}
nrow(dennys)
ncol(dennys)
view(dennys)
```

rows -> 1643
columns -> 6

Columns are address, city, state, zip, latitude, and longitude
Each row is a specific Dennys location

### Exercise 2

```{r}
nrow(laquinta)
ncol(laquinta)
view(laquinta)
```

rows -> 909
columns -> 6

The columns are the address city state zip latitude longitude
Each row of the data is a specific La Quinta location from around the world

### Exercise 3

La Quinta has locations outside the US including Mexico Canada and Columbia to name a few
There is no out of US Dennys

### Exercise 4

```{r}
view(laquinta)
view(dennys)
```

Use the filter function to find the out of country locations by the longitude and latitude columns

### Exercise 5
```{r}
dennys %>%
  filter(!(state %in% states$abbreviation))
```

There are no locations outside of the US for Dennys

### Exercise 6

```{r}
dennys %>%
  mutate(country = "United States")
```

### Exercise 7

```{r}
laquinta %>%
  filter(!(state %in% states$abbreviation))
```

Mexico
Canada
Columbia

### Exercise 8

```{r}
laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state == "MEX"           ~ "Mexico"
  ))
```

```{r}
laquinta <- laquinta %>%
  filter(state %in% state.abb)
```

### Exercise 9

```{r}
dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))
```

California has the most amount of Dennys
Delaware has the least amount of Dennys

```{r}
laquinta %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))
```

Texas has the most amount La Quintas 
Maine has the least amount of La Quintas
