---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Joseph White"
date: "10/16/22"
output: github_document
---


### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox)
```

```{r load-data, message = FALSE}
states <- read_csv("states.csv")
```

### Exercise 1

```{r}
nrow(dennys)
ncol(dennys)
view(dennys)
```
rows = 1643
columns = 6
columns -> address city state zip latitude longitude
each row is a specific location

### Exercise 2

```{r}
nrow(laquinta)
ncol(laquinta)
view(laquinta)
```
rows = 909
columns = 6
columns -> address city state zip latitude longitude
each row is a specific La Quinta location

### Exercise 3

No out of country dennys
There is out of country La Quinta... chile, columbia etc

### Exercise 4

```{r}
view(dennys)
view(laquinta)
```
we could use the filter function on the data to find out of country locations


### Exercise 5

```{r}
dennys %>%
  filter(!(state %in% states$abbreviation))
```
no locations outside the US for dennys

### Exercise 6
```{r}
dennys %>%
  mutate(country = "United States")
```

## Exercise 7
```{r}
laquinta %>%
  filter(!(state %in% states$abbreviation))
```
mexico
colombia
Canada
chile

##exercise 8
```{r}
laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state == "MEX" ~ "Mexico",
    state == "CHI" ~ "Chile"
  ))

```
```{r}
laquinta <- laquinta %>%
  filter(state %in% state.abb)
```


## exercise 9
```{r}
dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))
```
California has the most dennys, Delaware has the least

```{r}
laquinta %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))
```
Texas has the most Laquintas, Maine has the least

##Exerice 10
```{r}
dennys <- dennys %>%
  mutate(establishment = "Denny's")
laquinta <- laquinta %>%
  mutate(establishment = "La Quinta")

dn_lq <- bind_rows(dennys, laquinta)

ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```

