---
title: "Lab 05"
author: "Joseph White"
date: "10/19/22"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox)
```


## ex 1
```{r}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```

## ex 2
```{r}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```


##ex 3
```{r}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```

6 distances that need to be calcullated

## ex 4
```{r}
summary(dn_lq_ak)
nrow(dn_lq_ak)
```

There are 11 varaibles with 6 observations each
variables names: address.x, city.x, state, zip.x, longitude.x, latitude.x, address.y, city.y, zip.y, longitude.y, latitude.y

## exercise 5
we should use the mutate function?

#exercise 6
```{r}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```

```{r}
dn_lq_ak = dn_lq_ak %>% mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 4))
```

# exercise 7
```{r}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

# exercise 8
```{r}
df <- dn_lq_ak_mindist %>% select(closest)
boxplot(df, show.names=TRUE)
```

# exercise 9
```{r}
dn_nc <- dennys %>%
  filter(state == "NC")

lq_nc <- laquinta %>%
  filter(state == "NC")

dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")

dn_lq_nc = dn_lq_nc %>% mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 4))

dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

```{r}
df <- dn_lq_nc_mindist %>% select(closest)
boxplot(df, show.names=TRUE)
```

# exercise 10
```{r}
dn_tx <- dennys %>%
  filter(state == "TX")

lq_tx <- laquinta %>%
  filter(state == "TX")

dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")

dn_lq_tx = dn_lq_tx %>% mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 4))

dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

```{r}
df <- dn_lq_tx_mindist %>% select(closest)
boxplot(df, show.names=TRUE)
```


# exercise 11
```{r}
dn_me <- dennys %>%
  filter(state == "ME")

lq_me <- laquinta %>%
  filter(state == "ME")

dn_lq_me <- full_join(dn_me, lq_me, by = "state")

dn_lq_me = dn_lq_me %>% mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 4))

dn_lq_me_mindist <- dn_lq_me %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

```{r}
df <- dn_lq_me_mindist %>% select(closest)
boxplot(df, show.names=TRUE)
```

# exercise 12
Mitch Hedberg’s joke held true the most for Alaska. We decided so because the distribution of the closest distances is most consistently low and there are less outliers than Texas

