---
title: "money_in_US_politics"
author: "Sarah Kohls"
date: "2022-11-03"
output: html_document
---

##Packages
```{r}
library(tidyverse)
library(robotstxt)
library(rvest)
library(scales)
library(dsbox)
```

##Data collection
```{r}
paths_allowed("https://www.opensecrets.org")
```

##Scraping 2020 contributions
```{r}
url_2020 <- "https://www.opensecrets.org/political-action-committees-pacs/foreign-connected-pacs/2020"
```

##Read the webpage
```{r}
page <- read_html(url_2020)
```

##Scrape
```{r}
pac_2020 <- page %>%
  html_node(".DataTable-Partial") %>%
  html_table("td", header = TRUE, fill = TRUE) %>%
  as_tibble()
```

##Glipmse dataset
```{r}
glimpse(pac_2020)
```

##Number of observations
```{r}
nrow(pac_2020)
```

##Number of variables
```{r}
ncol(pac_2020)
```

##Separate country and parent
```{r}
pac_all_raw <- pac_all_raw %>%
  separate(country_parent, into = c("country", "parent"), sep = "/", extra = "merge")
```

##Create function to convert monetary values to numeric
```{r}
parse_currency <- function(x){
  x %>% 
    str_remove("\\$") %>%
    str_remove_all(",") %>%
    as.numeric()
}
```

##Use the function
```{r}
pac_all <- pac_all_raw %>%
  mutate(
    total = parse_currency(total),
    dems = parse_currency(dems),
    repubs = parse_currency(repubs)
  )
```

##Glimpse the cleaned data
```{r}
glimpse(pac_all)
```

##Data visualization
```{r}
pac_all %>%
  filter(country %in% c("Canada", "UK")) %>%
  group_by(country, year) %>%
  summarise(tot = sum(total), .groups = "drop") %>%
  ggplot(aes(x = year, y = tot, group = country, color = country)) +
  geom_line()
```

##Build plot to show contribution to US politics from UK-Connected PACs

##Filter
```{r}
pac_all %>%
  filter(country == "UK")
```

##Calculate total contributions
```{r}
pac_all %>%
  filter(country == "UK") %>%
  group_by(year) %>%
  summarise(
    Democrat = sum(dems),
    Republican = sum(repubs),
    .groups = "drop"
  )
```

##Pivoting
```{r}
pac_all %>%
  filter(country == "UK") %>%
  group_by(year) %>%
  summarise(
    Democrat = sum(dems),
    Republican = sum(repubs),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(Democrat, Republican), 
    names_to = "party", 
    values_to = "amount"
    )
```

##Save as new data frame
```{r}
pac_all_final <- pac_all %>%
  filter(country == "UK") %>%
  group_by(year) %>%
  summarise(
    Democrat = sum(dems),
    Republican = sum(repubs),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(Democrat, Republican), 
    names_to = "party", 
    values_to = "amount"
    )
```

##Plot
```{r}
pac_all_final %>%
  ggplot(mapping = aes(x = year, y = amount, color = party)) +
  geom_line()
```

##Change party colors
```{r}
pac_all_final %>%
  ggplot(mapping = aes(x = year, y = amount, color = party)) +
  geom_line() +
  scale_color_manual(values = c("blue", "red"))
```

##Dollar formatting
```{r}
pac_all_final %>%
  ggplot(mapping = aes(x = year, y = amount, color = party)) +
  geom_line() +
  scale_color_manual(values = c("blue", "red")) +
  scale_y_continuous(labels = label_dollar(scale = 0.000001, suffix = "M"))
```

##Descriptive labels
```{r}
pac_all_final %>%
  ggplot(mapping = aes(x = year, y = amount, color = party)) +
  geom_line() +
  scale_color_manual(values = c("blue", "red")) +
  scale_y_continuous(labels = label_dollar(scale = 0.000001, suffix = "M")) +
  labs(
    x = "Year",
    y = "Amount",
    color = "Party",
    title = "Contribution to US politics from UK-Connected PACs",
    subtitle = "By party, over time"
  ) 
```

##Minimal theme
```{r}
pac_all_final %>%
  ggplot(mapping = aes(x = year, y = amount, color = party)) +
  geom_line() +
  scale_color_manual(values = c("blue", "red")) +
  scale_y_continuous(labels = label_dollar(scale = 0.000001, suffix = "M")) +
  labs(
    x = "Year",
    y = "Amount",
    color = "Party",
    title = "Contribution to US politics from UK-Connected PACs",
    subtitle = "By party, over time"
  ) +
  theme_minimal()
```

##Plot is finished
##Finished intro to web scraping

