---
title: "NYC Flights"
author: "Sarah Kohls"
data: "2022-09-12"
output: html_notebook
---

Class play with NYC flights data

##Load Packages
```{r}
library(nycflights13)
library(tidyverse)
library(here)
library(janitor)
```

## Load flight data
```{r}
df1=flights
```

## Class work
```{r}
df2=df1 %>%
  filter(origin=="MIA",month==9)
```

```{r}
df3=df1 %>%
  filter(month==9,dest=="MIA")
```

```{r}
df4=df1 %>%
  filter(month==1,dest=="MIA")
```

```{r}
df5=df1 %>%
  filter(month==7,dest=="MIA")
```

```{r}
df6=df1 %>%
  filter(dest=="MIA")

df6 %>% ggplot(aes(month)) + geom_bar()
```

Analyzing the graph above we note:

* December has the most flights, perhaps due to holiday. 
* Feb is low, perhaps due to fewer days in Feb.
* Flight numbers seem to correlate with days in the month.

```{r}
df7=df1 %>%
  filter(dest %in% c("LAX","SAN","SFO","ATL","ORD","MIA"))

df7 %>% ggplot(aes(x=month,fill=dest)) + geom_bar()
```
```{r}
df7 %>% ggplot(aes(month,fill=dest)) + geom_bar()+facet_wrap(~dest)

```
Observations from the above graphs:

* San Diego is much smaller than the others.
* MIA is unusually flat.
* ORD is summer intensive (more traffic over summer)
* ORD, LAX, SFO patterns are similar: low in winter, high in summer.
* ATL is flat, perhaps it's a hub.
* February is low (fewer days) and winter.

```{r}
thisfile=here("Downloads","DL_SelectFields.zip")

df2019=read_csv(thisfile) %>% clean_names()
```

```{r}
df9=df2019 %>% filter(passengers>0,origin=="LGA")

df9 %>% ggplot(aes(month)) + geom_bar()
```

```{r}
df9 %>% ggplot(aes(month)) + geom_bar(aes(weight=departures_performed))
```
 Observations from the above plot: 
 
 * The months with the least amount of flights are January and February.
 * Other than Jan & Feb, most of the months have a similar number of flights.
 * There is also a relatively low number of flights in November and December.
 
```{r}
df9 %>% ggplot(aes(group=month,x=month, y=passengers)) + geom_boxplot()
```
 
```{r}
df9 %>% ggplot(aes(month,fill=carrier_name)) + geom_bar(aes(weight=departures_performed))
```
 
